#Reproducible Research: Peer Assessment 1

<p>Marius Florin RADU
<br>Cluj-Napoca, Cluj, ROMANIA
<br>mail: radu_marius_florin@yahoo.com</p>



## Loading and preprocessing the data


Read the data source:
```{r load_data}
activityDataFrame <- read.table(unz("activity.zip", "activity.csv"), 
                                colClasses = c("numeric", "Date", "numeric"),
                                header=T, quote="\"", sep=",")
head(activityDataFrame, 3)
summary(activityDataFrame)
```


## What is mean total number of steps taken per day?

```{r mean_of_steps_x_day}
# Use data.table to compute mean and median
library(data.table)
DT <- data.table(activityDataFrame, key = c("interval"))
# Calculate the mean
(mMean <- DT[, sum(steps), by = date][,list(Mean = mean(V1, na.rm = TRUE))])
# Calculate the median
(mMedian <- DT[, sum(steps), by = date][,list(Median = median(V1, na.rm = TRUE))])
```

The histogram of the total number of steps taken each day
```{r chart_steps_1, echo = TRUE, fig.width = 10, fig.height = 5}
library(ggplot2)
activityCompleteCases <- activityDataFrame[complete.cases(activityDataFrame$steps),] 
DT2 <- data.table(activityCompleteCases, key = c("interval"))
stepsDaily <- as.data.frame(DT2[, list(`Steps per Day`=sum(steps)), by = date])
# Histogram
hist1 <- qplot(`Steps per Day`, data=stepsDaily, geom="histogram") +
        geom_histogram(aes(fill = ..count..)) +
        xlab('Number of Steps per Day') +
        ggtitle('Histogram of the total number of steps taken each day') +
        theme(plot.title = element_text(lineheight=.8, face="bold"))
hist1
```

The __mean__ number of steps per day is: __`r as.character(round(mMean,0))`__

The __median__ number of steps per day is: __`r as.character(round(mMedian,0))`__


## What is the average daily activity pattern?

```{r chart_steps_2, echo = TRUE, fig.width = 13, fig.height = 6}
library(ggplot2)
dataFrameforTimeSeries <- DT[, list( `Average # of Steps` = mean(steps, na.rm = TRUE) ), by = interval]
p <- ggplot(dataFrameforTimeSeries , aes(x=interval, y=`Average # of Steps`)) + 
     geom_line(colour = "dark blue", size=0.8) + 
     theme(axis.text.x = element_text(angle = 90, hjust = 1))

p
```

```{r max_steps_is, echo = TRUE}
(maxSumSteps<-DT[, list( sumSteps = sum(steps, na.rm = TRUE) ), by = interval][, .SD[which.max(sumSteps)]])
theInterval <- maxSumSteps[[1]]
theMaxSum <-maxSumSteps [[2]]
```

The 5-minute interval which contains the maximum number of steps is __`r paste(theInterval, "th", sep="")`__ interval, and the corespondent number of steps is __`r as.character(theMaxSum)`__ 

## Imputing missing values

```{r}
isNA <- dim(activityDataFrame[!complete.cases(activityDataFrame),])[1]
```
The total number of rows with *NAs* is __`r isNA`__


**To replace missing values of steps variable in following rows we use the mean computed for corespondent week of the record.**
```{r}
head(activityDataFrame, 3)
activityDataFrame02 <- activityDataFrame
activityDataFrame02$Week <- format(activityDataFrame02$date,"%Y-%U")
stepsMeanByWeek <- aggregate(steps ~ Week, activityDataFrame02, FUN = mean, na.rm = TRUE)
for( i in seq( dim(activityDataFrame02)[1] )){
        if( is.na(activityDataFrame02[i , c("steps")]) ){
                activityDataFrame02[i, c("steps")] <- stepsMeanByWeek[which(stepsMeanByWeek$Week == activityDataFrame02[i, c("Week")] ),c("steps")]
        }
}
activityDataFrame02 <- activityDataFrame02[,-c(4)]
```


The result is a new dataset that is equal to the original dataset but with the missing data filled in:
```{r}
head(activityDataFrame02, 3)
summary(activityDataFrame02)
```


```{r mean_comparison}
# Use data.table to compute mean and median
library(data.table)
DTII <- data.table(activityDataFrame02, key = c("interval"))
# Calculate the mean
(mMeanII <- DTII[, sum(steps), by = date][,list(Mean = mean(V1, na.rm = TRUE))])
# Calculate the median
(mMedianII <- DTII[, sum(steps), by = date][,list(Median = median(V1, na.rm = TRUE))])
```

The histogram of the total number of steps taken each day
```{r chart_comparisons, echo = TRUE, fig.width = 10, fig.height = 5}
library(ggplot2)
activityCompleteCasesII <- activityDataFrame02[complete.cases(activityDataFrame02$steps),] 
DT2II <- data.table(activityCompleteCasesII, key = c("interval"))
stepsDailyII <- as.data.frame(DT2II[, list(`Steps per Day`=sum(steps)), by = date])
# Histogram
hist2 <- qplot(`Steps per Day`, data=stepsDailyII, geom="histogram") +
        geom_histogram(aes(fill = ..count..)) +
        xlab('Number of Steps per Day') +
        ggtitle('Histogram of the total number of steps taken each day after NA replacement') +
        theme(plot.title = element_text(lineheight=.8, face="bold")) + 
        scale_fill_gradient("Count", low = "dark blue", high = "red")
hist2
```

The __mean__ number of steps per day after **NAs** replacement is: __`r as.character(round(mMeanII,0))`__ and it is not statistically different from the __mean__ before replachement which is __`r as.character(round(mMean,0))`__

There is bellow the **t-test** evaluation 
```{r t_test_means, echo =TRUE}
activityDataFrame02$old_steps <- activityDataFrame$steps
(t.test(activityDataFrame02$steps,activityDataFrame02$old_steps))
```

The __median__ number of steps per day after **NAs** replacement is: __`r as.character(round(mMedianII,0))`__ and it is not statistically different from the __median__ before replachement which is __`r as.character(round(mMedian,0))`__

There is bellow the **Mann-Whitney-Wilcoxon test** evaluation 
```{r mann_w_median, echo =TRUE}
(wilcox.test(activityDataFrame02$steps,activityDataFrame02$old_steps))
```


## Are there differences in activity patterns between weekdays and weekends?

```{r plot_by_weekday}
# Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
#Decimal Weekday (0=Sunday)
activityDataFrame03 <- activityDataFrame
activityDataFrame03$DayOfWeek <- as.factor(ifelse(format(activityDataFrame$date,"%w") %in% c(0,6), 0, 1))
levels(activityDataFrame03$DayOfWeek) <- c("weekend", "weekday")
table(activityDataFrame03$DayOfWeek)

activityCompleteCasesIII <- activityDataFrame03[complete.cases(activityDataFrame03$steps),] 
head(activityCompleteCasesIII, 3)
# Histogram
tsPlot2 <- ggplot(activityCompleteCasesIII, aes(interval, steps)) +  
        ylab('Number of Steps') +
        ggtitle('# steps taken each interval by day of the week') +
        stat_summary(fun.y = "mean", geom="line", col = "red", size=0.8) + 
        facet_grid(DayOfWeek ~ .) +
        theme(axis.text.x = element_text(angle = 90, hjust = 1))
tsPlot2
```

**Yes**, there are differences between weekdays and weekends. 
