# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data

Read the data source:
```{r load_data}
activityDataFrame <- read.csv(unz("activity.zip", "activity.csv"), colClasses = c("numeric", "Date", "numeric"))
summary(activityDataFrame)
head(activityDataFrame, 3)
```

## What is mean total number of steps taken per day?

```{r mean_of_steps_x_day}
# Use data.table to compute mean and median
library(data.table)
DT <- data.table(activityDataFrame, key = c("interval"))
# Calculate mean and median
(mm <- DT[, sum(steps), by = date][,list(Mean = mean(V1, na.rm = TRUE), Median = median(V1, na.rm = TRUE))])
```

_The __mean__ number of steps per day is: `r round(as.integer(mm[[1]]),0)`

The __median__ number of steps per day is: __`r round(mm[[2]],0)`__

## What is the average daily activity pattern?

```{r echo = TRUE}
library(ggplot2)
dataFrameforTimeSeries <- DT[, list( `Average # of Steps` = mean(steps, na.rm = TRUE) ), by = interval]

dataFrameforTimeSeries 
str(dataFrameforTimeSeries$interval)
lvl <- 
p <- ggplot(dataFrameforTimeSeries , aes(x=interval, y=`Average # of Steps`)) + 
     geom_line(colour = "dark blue", size=0.8) + 
     theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
     scale_x_discrete(limits=lvl)

p
```

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
